---
title: "article-1"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Example

This is a basic example which shows you a basic workflow of the package:

## Load necessary packages

We first load R packages required for the example here.
```{r setup, message = FALSE}
library(patchwork)
library(ggtree)

# spatialCCC package
library(spatialCCC)
```

## Ligand-Receptor Database

### Load built-in LRdb

This LRdb is downloaded from CellTalkDB: [http://tcm.zju.edu.cn/celltalkdb/download.php].  You can see the detail by:

``` r
?LRdb_human
?LRdb_mouse
```

We also created getter functions, get_LRdb() and get_LRdb_small(), to retrieve the database.

```{r LRdb-loading}
set.seed(100)

LRdb_m <- 
  get_LRdb("mouse", n_samples = 100)

LRdb_m %>% 
  dplyr::arrange(ligand_gene_symbol, receptor_gene_symbol)
```

## Load Visium spatial transcriptomic data

```{r visium-data-loading}
data_dir <- spatialCCC_example("example")

spe_brain <-
  SpatialExperiment::read10xVisium(samples = data_dir,
                                   type = "HDF5",
                                   data = "filtered")

# to keep track of cell IDs
# spe_brain[["cell_id"]] <- colnames(spe_brain)

# Log-Normalize
spe_brain <- scater::logNormCounts(spe_brain) 
```


## Cell cluster data

Now, let's read cell cluster data, obtained using [GraphST](https://deepst-tutorials.readthedocs.io/).
```{r}
cell_clusters <-
  read.csv(file.path(data_dir, "outs", "graphST.csv"), row.names = 1)

# Make sure the rows of cell_clusters are in the same order of spe_brain.
cell_clusters <- cell_clusters[colnames(spe_brain), ]

cluster_ids <- colnames(cell_clusters)
cluster_ids <- cluster_ids[grep("mclust_", cluster_ids)]

for (cid in cluster_ids) {
  spe_brain[[cid]] <- factor(cell_clusters[[cid]])
}
```


```{r}
plot_spatial_feature(spe = spe_brain, feature = "mclust_8")
```


## Compute Cell-Cell Communications over ligand-receptor pairs

```{r compute-spatial-ccc, message=FALSE, warning=FALSE}
tictoc::tic()

ccc_graph_list <-
  compute_spatial_ccc_graph_list(spe = spe_brain,
                                 assay_name = "logcounts",
                                 LRdb = LRdb_m,
                                 workers = 8)

tictoc::toc()
```


```{r}
ccc_graph_list <-
  ccc_graph_list %>%
  purrr::map(function(ccog) {
    ccog %>%
      add_node_annot_to_edges(c("mclust_8", "mclust_9", "mclust_10"))
  })
```


```{r}
ccc_tbl <-
  ccc_graph_list %>%
  purrr::map(function(ccog) {
    ccog %>%
      tidygraph::activate("edges") %>%
      as_tibble()
  }) %>%
  dplyr::bind_rows(.id = "LR")
```





### Summarize cell-cell communication between cell clusters
```{r}
ccc_tbl_between_clusters <-
  ccc_tbl %>%
  dplyr::group_by(mclust_8_src, mclust_8_dst, LR) %>%
  dplyr::summarise(n = dplyr::n(),
                   LRscore.sum = sum(LRscore),
                   .groups = "drop")

ccc_tbl_between_clusters %>%
  dplyr::arrange(desc(n))
```

```{r}
ccc_tbl_between_clusters %>%
  dplyr::filter(mclust_8_dst != mclust_8_src) %>%
  dplyr::arrange(desc(n))
```



### Convert CCC table to CCC graph

summarize_ccc_graph_metrics() summarize those graph metrics for each LR pair.

```{r summarize-ccc-graph-metrics-graph}
tictoc::tic()

ccc_graph_metrics_summary_df <-
  summarize_ccc_graph_metrics(ccc_graph_list)

tictoc::toc()
```



```{r inspect-ccc-graph-metrics-summary}
ccc_graph_metrics_summary_df %>%
  dplyr::arrange(graph_component_count)
```


summarize_ccc_graph_metrics(..., level = "group") summarizes the metrics for each subgraph (group) in CCC graph (LR)

```{r summarise-ccc-graph-metrics-group}
tictoc::tic()

ccc_graph_group_metrics_summary_df <-
  summarize_ccc_graph_metrics(ccc_graph_list, level = "group")

tictoc::toc()
```


```{r inspect-ccc-graph-metrics-group-summary}
ccc_graph_group_metrics_summary_df 
```



## Visualization

```{r example-LR, fig.width=8}
LR_of_interest <- "Pdgfb_Pdgfra"
```


```{r}
ccc_graph_list[[LR_of_interest]] %>%
  tidygraph::activate(edges) %>%
  tibble::as_tibble()
```



### spatial CCC graph plot with tissue image

```{r}
gp_spccc <-
  plot_spatial_ccc_graph(
    ccc_graph = ccc_graph_list[[LR_of_interest]],
    tissue_img = SpatialExperiment::imgRaster(spe_brain),
    node_color = "mclust_8",
    node_size = 1,
    node_alpha = 0.5,
    edge_color = "group",
    # clip = TRUE,
    which_on_top = "edge"
  )

gp_spccc_0 <-
  plot_spatial_feature(spe = spe_brain,
                       feature = "mclust_8")
```

```{r plot-spatial-ccc-graph-tissue-image, warning=FALSE}
wrap_plots(gp_spccc, gp_spccc_0, ncol = 2)
```



```{r}
ccc_graph_temp <-
  ccc_graph_list[[LR_of_interest]] %>%
  tidygraph::activate("edges") %>%
  tidygraph::filter(mclust_8_src != mclust_8_dst) %>%
  tidy_up_ccc_graph()

cells_of_interest <-
  ccc_graph_temp %>%
  tidygraph::activate("nodes") %>%
  dplyr::pull("name")
```

```{r}
gp_spccc <-
  plot_spatial_ccc_graph(
    ccc_graph = ccc_graph_temp,
    tissue_img = SpatialExperiment::imgRaster(spe_brain),
    node_color = "n.inflow",
    node_size = 1.25,
    node_alpha = 1,
    edge_color = "group",
    show_arrow = TRUE,
    # clip = TRUE,
    which_on_top = "node"
  )

gp_spccc_0 <-
  plot_spatial_ccc_graph(
    ccc_graph = ccc_graph_list[[LR_of_interest]],
    tissue_img = SpatialExperiment::imgRaster(spe_brain),
    image_alpha = 0,
    cells_of_interest = cells_of_interest,
    edges_expanded_to_group = FALSE,
    node_color = "mclust_8",
    node_size = 1.25,
    node_alpha = 0.5,
    edge_color = "group",
    show_arrow = TRUE,
    # clip = TRUE,
    which_on_top = "node"
  )

gp_spccc_1 <-
  plot_spatial_feature(spe = spe_brain,
                       feature = "mclust_8",
                       cells_of_interest = cells_of_interest)
```

```{r plot-spatial-ccc-graph-tissue-overlapped-cell-cluster, warning=FALSE, fig.width=9, fig.height=6}
wrap_plots(gp_spccc, gp_spccc_1, ncol = 2)
```

```{r, fig.width=8}
gp_spccc_0
```


### spatial CCC graph plot without tissue image

In this case, graph layout can be "spatial" which keeps the original spatial locations, or other graph layout algorithm supported by igraph package. 


```{r}
gp_spccc <-
  plot_spatial_ccc_graph(
    ccc_graph = ccc_graph_list[[LR_of_interest]],
    graph_layout = "spatial",
    node_color =  "mclust_8",
    node_size = 1,
    edge_color = "group_diameter",
    clip = TRUE,
    # ghost_img = TRUE,
    which_on_top = "edge"
  )

gp_spccc_0 <-
  plot_spatial_ccc_graph(
    ccc_graph =
      ccc_graph_list[[LR_of_interest]],
    graph_layout = "spatial",
    node_color = "mclust_8",
    node_size = 1,
    edge_color = "group_diameter",
    clip = TRUE,
    # ghost_img = TRUE,
    which_on_top = "node"
  )
```

```{r plot-spatial-ccc-graph-spatial, warning=FALSE, fig.width=10}
wrap_plots(gp_spccc, gp_spccc_0, ncol = 2, guides = "collect")
```


Below uses "auto" layout ("kk" spring layout).

```{r plot-spatial-ccc-graph-kk-layout, warning=FALSE}
plot_spatial_ccc_graph(ccc_graph = ccc_graph_list[[LR_of_interest]],
                       # tissue_img = SpatialExperiment::imgRaster(spe_brain),
                       node_color = "group",
                       node_size = 0.1,
                       edge_color = "group_diameter",
                       edge_width = 0.1,
                       which_on_top = "edge")
```

In this case, below is "stress" layout.

```{r plot-spatial-ccc-graph-stress-layout, warning=FALSE}
plot_spatial_ccc_graph(ccc_graph = ccc_graph_list[[LR_of_interest]],
                       # tissue_img = SpatialExperiment::imgRaster(spe_brain),
                       graph_layout = "stress",
                       node_color = "group",
                       edge_color = "group_diameter",
                       edge_width = 0.25,
                       which_on_top = "edge")
```



## Cell-overlap distance
```{r compute-cell-overlap-distance}
tictoc::tic()

cell_overlap_dist <-
  dist_cell_overlap_ccc_tbl(ccc_tbl)

tictoc::toc()
```

```{r compute-cell-overlap-distance-long-format}
tictoc::tic()

cell_overlap_lf <-
  lf_cell_overlap_ccc_tbl(ccc_tbl)

tictoc::toc()
```

```{r select-LR-with-high-cell-overlap}
tictoc::tic()

LRs_high_cell_overlap <-
  cell_overlap_lf %>% 
  dplyr::filter(d < 1) %>%
  dplyr::select(lr1, lr2) %>%
  unlist() %>% unique()

tictoc::toc()
```

```{r subset-high-cell-overlap-LR}
tictoc::tic()

high_cell_overlap_dist <-
  cell_overlap_dist[LRs_high_cell_overlap, LRs_high_cell_overlap]

tictoc::toc()
```

```{r subset-high-cell-overlap-LR-2}
tictoc::tic()

high_cell_overlap_dist2 <-
  cell_overlap_lf %>%
  dplyr::filter(d < 1) %>%
  dplyr::select(lr1, lr2, d) %>%
  lf_to_dist()

tictoc::toc()
```

```{r summarize-ccc-table-by-LR}
LR_ccc_summary_tbl <-
  ccc_tbl %>% 
  dplyr::pull(LR) %>%
  table() %>%
  tibble::as_tibble() %>%
  dplyr::rename("LR" = ".") %>%
  dplyr::arrange(desc(n)) %>%
  dplyr::left_join(
    ccc_tbl %>% 
      dplyr::select(LR, ligand, receptor) %>%
      dplyr::distinct(),
    by = "LR"
  )
```


```{r clustering-LR-by-cell-overlap}
hclus.res <- fastcluster::hclust(as.dist(high_cell_overlap_dist),
                                 method = "complete")
```


```{r LR-similarity-flat-dendrogram, fig.width=8, fig.height=2}
#| warning: false
ape::as.phylo(hclus.res) %>%
  ggtree(layout="dendrogram") %<+% LR_ccc_summary_tbl +
  aes(color=receptor) +
  theme(legend.position = "none") +
  geom_tippoint(aes(size=n), alpha=0.5) +
  geom_tiplab(size=2, offset=-0.15) +
  xlim_tree(3)
```


```{r LR-similarity-circular-dendrogram, fig.width=5, fig.height=5}
#| warning: false
ape::as.phylo(hclus.res) %>%
  ggtree(layout = "circular") %<+% LR_ccc_summary_tbl +
  aes(color=receptor) +
  # aes(color=ligand) +
  theme(legend.position = "none") +
  geom_tippoint(aes(size=n), alpha=0.5) +
  geom_tiplab(size=2, offset=0.05)
```

