---
title: "spatialCCC"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(spatialCCC)
```


```{r}
library(dplyr)
library(ggplot2)
library(patchwork)
```


  * Then, load built-in LR database.
```{r LRdb-loading}
set.seed(100)

LRdb_m <- 
  get_LRdb("mouse", n_samples = 50) 

LRdb_full <-
  get_LRdb("mouse")

LRdb <- LRdb_m; n_workers = 1
#LRdb <- LRdb_full; n_workers = 8
```


  * Load an example Visium spatial transcriptomic data
```{r visium-data-loading}
# absolute path to "example" spatial transcriptomic data
data_dir <- spatialCCC_example("example")

spe_brain <-
  SpatialExperiment::read10xVisium(samples = data_dir,
                                   type = "HDF5",
                                   data = "filtered")

# Log-Normalize
spe_brain <- scater::logNormCounts(spe_brain)
```



## Cell cluster data

Now, let's read cell cluster data, obtained using [GraphST](https://deepst-tutorials.readthedocs.io/).
```{r}
cell_clusters <-
  read.csv(file.path(data_dir, "outs", "graphST.csv"), row.names = 1)

# Make sure the rows of cell_clusters are in the same order of spe_brain.
cell_clusters <- cell_clusters[colnames(spe_brain), ]

cluster_ids <- colnames(cell_clusters)
cluster_ids <- cluster_ids[grep("mclust_", cluster_ids)]

for (cid in cluster_ids) {
  spe_brain[[cid]] <- factor(cell_clusters[[cid]])
}
```


```{r}
plot_spatial_feature(spe = spe_brain, feature = "mclust_8")
```



### Cell-cell communication analysis

  * Compute Cell-Cell Communications over ligand-receptor pairs
```{r compute-spatial-ccc, message=FALSE}
#future::plan(future::multisession, workers = 8)

ccc_graph_list <-
  compute_spatial_ccc_graph_list(spe = spe_brain,
                                 # uncomment the line below when testing
                                 # comment the line before uploading to github
                                 workers = n_workers,
                                 assay_name = "logcounts",
                                 LRdb = LRdb)

#future::plan(future::sequential)
```



```{r}
ccc_graph_list <-
  ccc_graph_list %>%
  purrr::map(function(ccog) {
    ccog %>%
      transfer_node_annots_to_edges(c("mclust_8", "mclust_9", "mclust_10"))
  })
```



```{r}
ccc_graph_tbl <-
  ccc_graph_list %>%
  purrr::map(function(cg) {
    cg %>%
      activate("edges") %>%
      as_tibble()
  }) %>%
  bind_rows(.id = "LR")
```


```{r}
ccc_graph_tbl %>%
  group_by(LR) %>% 
  summarise(n = n()) %>%
  arrange(desc(n))
```


```{r, fig.width=8}
LR_of_interest <- "App_Lrp1"

ccc_graph_interfacing <-
  ccc_graph_list[[LR_of_interest]] %>%
  tidygraph::activate("edges") %>%
  tidygraph::filter(mclust_8_src != mclust_8_dst) %>%
  tidy_up_ccc_graph()


plot_spatial_ccc_graph(
  ccc_graph = ccc_graph_interfacing,
  tissue_img = SpatialExperiment::imgRaster(spe_brain),
  graph_layout = "spatial",
  edges_expanded_to_group = TRUE,
  node_color = "mclust_8",
  node_size = 2.6,
  edge_color = "LRscore",
  show_arrow = TRUE,
  clip = TRUE,
  which_on_top = "node"
) 
```



# Build cluster-cluster interface
```{r}
c2c_dist <-
  calc_spot_dist(spe = spe_brain)

c2c_dist[["mclust_8_src"]] <- cell_clusters[c2c_dist$src, "mclust_8"]
c2c_dist[["mclust_8_dst"]] <- cell_clusters[c2c_dist$dst, "mclust_8"]

c2c_n <-
  c2c_dist %>%
  group_by(mclust_8_src, mclust_8_dst) %>%
  summarize(N = n(), .groups = "drop") %>%
  mutate(mclust_8_src = factor(mclust_8_src), mclust_8_dst = factor(mclust_8_dst))
```



```{r}
ccc_tbl <-
  ccc_graph_list %>%
  purrr::map(function(ccog) {
    ccog %>%
      tidygraph::activate("edges") %>%
      as_tibble()
  }) %>%
  dplyr::bind_rows(.id = "LR")
```


### Summarize cell-cell communication between cell clusters
```{r}
ccc_tbl_between_clusters <-
  ccc_tbl %>%
  dplyr::group_by(mclust_8_src, mclust_8_dst, LR) %>%
  dplyr::summarise(n = dplyr::n(),
                   LRscore.sum = sum(LRscore),
                   LRscore.median = median(LRscore),
                   .groups = "drop") %>%
  dplyr::left_join(c2c_n, by = c("mclust_8_src", "mclust_8_dst")) %>%
  dplyr::mutate(r = n/N)
```


```{r}
ccc_tbl_between_clusters %>%
  dplyr::filter(N > 10) %>%
  dplyr::group_by(mclust_8_src, mclust_8_dst) %>%
  dplyr::arrange(desc(r)) %>%
  dplyr::slice_head(n = 2) %>%
  dplyr::ungroup() 
```


```{r}
ccc_tbl_between_clusters %>%
  dplyr::filter(N > 10, r > 0.5) %>%
  dplyr::arrange(mclust_8_src, mclust_8_dst, desc(r)) 
```


```{r}
LR_of_interest <- "App_Lrp1"

ccc_graph_temp <-
  ccc_graph_list[[LR_of_interest]] %>%
  tidygraph::activate("edges") %>%
  tidygraph::filter(mclust_8_src == 1, mclust_8_dst == 6) %>%
  tidy_up_ccc_graph()

cells_of_interest <-
  ccc_graph_temp %>%
  tidygraph::activate("nodes") %>%
  dplyr::pull("name")
```


```{r}
gp_spccc <-
  plot_spatial_ccc_graph(
    ccc_graph = ccc_graph_temp,
    tissue_img = SpatialExperiment::imgRaster(spe_brain),
    node_color = "mclust_8",
    # node_size = 1.25,
    node_alpha = 1,
    edge_color = "LRscore",
    edge_width = 1.5,
    show_arrow = TRUE,
    clip = TRUE,
    which_on_top = "node"
  )

gp_spccc
```


```{r}
LRs <- unique(ccc_tbl_between_clusters$LR)

names(LRs) <- LRs

xx <-
  LRs %>%
  purrr::map(function(lr) {
    x <- ccc_tbl_between_clusters %>% filter(LR == lr)
    x %>%
      arrange(desc(LRscore.median)) %>%
      filter(n > 10) %>%
      filter(mclust_8_src != mclust_8_dst) %>%
      head(n = 2)
  }) %>%
  bind_rows(.id = "LR")
  
xx
```


```{r}
xxx <-
  purrr::map2(rep(1:8, times = 8), rep(1:8, each = 8), function(c1, c2) {
    x <- ccc_tbl_between_clusters %>% filter(mclust_8_src == c1, mclust_8_dst == c2)
    x %>%
      arrange(desc(LRscore.median)) %>%
      filter(n > 10) %>%
      filter(mclust_8_src != mclust_8_dst) %>%
      filter(LR != "App_Lrp1") %>%
      head(n = 1)
  }) %>%
  bind_rows()
  
xxx
```



  * Summarize spatial CCC graphs by LR
```{r}
ccc_graph_summary_tbl <-
  ccc_graph_list %>%
  to_LR_summary_tbl() %>%
  arrange(desc(pRank))

ccc_graph_summary_tbl
```




```{r}
top_LR <- 
  ccc_graph_summary_tbl %>%
  pull(LR) %>%
  head(n = 1)

bottom_LR <-
  ccc_graph_summary_tbl %>%
  pull(LR) %>%
  tail(n = 1)
```


```{r}
gp_ccc_graph_top_LR <-
  plot_spatial_ccc_graph(ccc_graph_list[[top_LR]],
                         graph_layout = "spatial",
                         edge_color = "LRscore",
                         node_color = "LRscore.sum.src") +
  ggtitle(top_LR)

gp_ccc_graph_bottom_LR <-
  plot_spatial_ccc_graph(ccc_graph_list[[bottom_LR]],
                         graph_layout = "spatial",
                         edge_color = "LRscore",
                         node_color = "LRscore.sum.src") +
  ggtitle(bottom_LR)
```


```{r, fig.width=10, fig.height=5}
wrap_plots(gp_ccc_graph_top_LR, gp_ccc_graph_bottom_LR)
```



```{r}
ccog_1_minus_2 <-
  ccc_graph_subtract(ccc_graph_list[[top_LR]],
                     ccc_graph_list[[bottom_LR]])

ccog_2_minus_1 <-
  ccc_graph_subtract(ccc_graph_list[[bottom_LR]],
                     ccc_graph_list[[top_LR]])
ccog_diff <-
  ccc_graph_diff(ccc_graph_list[[top_LR]],
                 ccc_graph_list[[bottom_LR]])
```



```{r}
ccog_union <-
  ccc_graph_union(ccc_graph_list[[top_LR]],
                  ccc_graph_list[[bottom_LR]])

ccog_intersect <-
  ccc_graph_intersect(ccc_graph_list[[top_LR]],
                      ccc_graph_list[[bottom_LR]])
```



```{r}
gp_list <-
  list(ccog_1_minus_2, ccog_2_minus_1, ccog_diff) %>%
  purrr::map(function(ccog) {
    plot_spatial_ccc_graph(ccog,
                           graph_layout = "spatial")
  })
```


```{r, fig.width=15, fig.height=5}
wrap_plots(gp_list)
```



```{r}
gp_list <-
  list(intersect = ccog_intersect, union = ccog_union) %>%
  purrr::map(function(ccog) {
    plot_spatial_ccc_graph(ccog,
                           graph_layout = "spatial")
  })
```


```{r, fig.width=10, fig.height=5}
wrap_plots(gp_list)
```








```{r}
gp_list <-
  list(ccog_1_minus_2, ccog_2_minus_1, ccog_diff) %>%
  purrr::map(function(ccog) {
    plot_spatial_ccc_graph(ccog,
                           graph_layout = "spatial",
                           tissue_img = SpatialExperiment::imgRaster(spe_brain))
  })
```


```{r, fig.width=15, fig.height=5}
wrap_plots(gp_list)
```




```{r}
gp_list <-
  list(intersect = ccog_intersect, union = ccog_union) %>%
  purrr::map(function(ccog) {
    plot_spatial_ccc_graph(ccog,
                           graph_layout = "spatial",
                           tissue_img = SpatialExperiment::imgRaster(spe_brain))
  })
```


```{r, fig.width=10, fig.height=5}
wrap_plots(gp_list)
```





```{r}
ccog_flattened <-
  flatten_ccc_graph_list(ccc_graph_ls = ccc_graph_list)
```


```{r, fig.height=6}
plot_spatial_ccc_graph(
  ccog_flattened,
  tissue_img = SpatialExperiment::imgRaster(spe_brain),
  graph_layout = "spatial",
  node_color = "LRscore.sum.src",
  edge_color = "LRscore",
  # node_size = 1.6,
  node_alpha = 0,
  which_on_top = "edge"
) 
```

```{r}
# top_x <-
#   extract_ccc_graph_metrics(ccc_graph = ccc_graph_list[[top_LR]], level = "group")
# bottom_x <-
#   extract_ccc_graph_metrics(ccc_graph = ccc_graph_list[[bottom_LR]], level = "group")
# top_x <-
#   summarize_ccc_graph_metrics(ccc_graph = ccc_graph_list[top_LR], level = "group")
# bottom_x <-
#   summarize_ccc_graph_metrics(ccc_graph = ccc_graph_list[bottom_LR], level = "group")
top_x <-
  extract_ccc_graph_metrics(ccc_graph = ccc_graph_list[[top_LR]], level = "group") %>%
  select(-contains("LR"))
bottom_x <-
  extract_ccc_graph_metrics(ccc_graph = ccc_graph_list[[bottom_LR]], level = "group") %>%
  select(-any_of("LR"))

# compute 
xz <-
  purrr::map(setNames(nm = colnames(top_x)), function(i) {
    x <-
      ks.test(top_x[[i]], bottom_x[[i]])
    x[c("statistic", "p.value", "alternative", "method", "exact")] %>%
      as_tibble()
  }) %>%
  bind_rows(.id = "name")
```


```{r}
ccc_graph_metrics_tbl <-
  ccc_graph_list %>%
  purrr::map(function(cg) {
    extract_ccc_graph_metrics(cg, level = "group")
  }) %>%
  bind_rows(.id = "LR")
```

